Reliable feature point matching without the use of geometric constraints

### 2013-06-06 text/clustering.tex
Made the paper better still

\documentclass[12pt,journal,compsoc]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
\usepackage[font={small}]{caption}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\title{Reliable feature point matching without the use of geometric 
constraints}
\author{Jonas Arnfred and what to put here exactly?}

@@

containing view port changes and a simple method to generate a large 

@@

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing every 
feature point of one image with all feature points of the other and 
finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. By ranking the 
scores by their uniqueness and picking the $n$ best we get a set of 
correspondences that are distinctly matched across the two images. This 
method will be referred to as \emph{SIFT} in this paper.

In turn the two methods proposed in this paper are inspired by a simple 
but novel idea. If we have two images and a given feature point in the 
first image is better matched with other feature points from the 
\emph{same} image than points in the other image, then any matches of 
this feature point to points in the other image is deemed unreliable and 
discarded.  This approach carries no implicit assumptions about the 
geometric consistency of matches and can as such easily be extended with 
other geometric solutions when appropriate.

@@

@@

@@

		\begin{subfigure}[t]{0.30\textwidth}

@@

			\caption{SIFT}

@@

		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.30\textwidth}

@@

		~ %add desired spacing between images, e. g. ~, \quad, \qquad
		  %(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.30\textwidth}

@@

		\end{subfigure}

@@

matches that are found within the same image. }
	\label{fig:comparemirror}
\end{figure*}
Figure \ref{fig:comparemirror} shows an example of the result of 
\emph{MM}. Notice how the by the bottom of the image admits 
correspondences that can conveniently be removed as done in sub figure 
\ref{fig:without} where only the actual correspondences returned by the 
algorithm are shown.
\subsection{Mirror Match with Clustering (\emph{MMC})}
As opposed to \emph{MM}, \emph{MMC} diverges from traditional 
non-geometric feature matching by clustering the feature points by 
similarity. This process yields partitions of fairly similar feature 
points that we can match using the same approach as \emph{MM}. 
%Before we introduce the implementation details we will go over the 
%problem of graph clustering and how it relates to feature matching.
%\subsubsection{Graph Clustering}
%Given a set of feature points from two images, $im_1$ and $im_2$: $F_k = 
%\{k_i, k_j$ for $k_i \in im_1, k_j \in im_2\}$ as well as a matching 
%function $M(k_i, k_j) \rightarrow \mathbb{R}$ that takes two feature 
%points in an image and returns their matching score, we can define a 
%matrix $A$ where each element $A_ij = M(k_i, k_j)$. A can be interpreted 
%as the \emph{adjecency matrix} of the fully connected graph where each 
%vertex corresponds to a keypoint and the edge between two vertices has a 
%weight equal to the distance between the two corresponding keypoints.
%This representation reduces the problem of partitioning the keypoints 
%into groups to that of graph clustering or community structure depending 
%on the context. In the literature there are various ways of clustering a 
%graph according to different measures of what constitutes an optimal 
%partitioning. Traditionally the most used clustering algorithms have 
%been K-means and spectral clustering, but in recent years a host of new 
%algorithms have been proposed based on both Newman's concept of graph 
%modularity\footnote{Introduced in \cite{girvan2002}, discussed in 
%\cite{brandes2007} and used in \cite{blondel2008} as well as others} as 
%well as information theoretical measures\footnote{See for example 
%\cite{rosvall2008}} and the Potts spin model from physics\footnote{Used 
%in \cite{ronhovde2009}} just to mention a few approaches. Many of the 
%new algorithms differ from K-means clustering and Spectral clustering in 
%that they don't require the number of expected to clusters to be 
%specified beforehand\footnote{Among the aforementioned methods, this is 
%true for \cite{blondel2008} and \cite{rosvall2008}}.  Furthermore, on 
%tests done using randomly generated graphs with a known partitioning 
%\cite{blondel2008}, \cite{rosvall2008} and \cite{ronhovde2009} perform 
%markedly better than spectral clustering and 
%K-means\cite{lancichinetti2009}.
%The performance of clustering algorithms is a complicated issue, since 
%an optimal clustering given the same graph can vary depending on the 
%application. Spectral clustering for example will usually return a 
%partitioning where each partition is roughly equal in size\footnote{As 
%mentioned in \cite{von2007}} while the Louvain 
%clustering\cite{blondel2008} might return partitions of very uneven 
%size, even if the modularity measurement has been shown to penalize very 
%small clusters\cite{brandes2007}. Both behaviours can be beneficial 
%depending on the application, but when clustering feature points, 
%maintaining clusters of an even size usually means that some clusters 
%will be '\emph{catch-all}' clusters where the feature points that don't 
%fit anywhere else are grouped together. The necessity of specifying the 
%amount of partitions in for example Spectral clustering or Pott's model 
%clustering further exacerbates the issue since smaller partitions are 
%then combined into one to achieve the right amount of partitions.
We use the Louvain Method\cite{blondel2008} for clustering feature 
points since it's relatively fast and performs 
well\cite{lancichinetti2009}, doesn't require 
parameters\cite{blondel2008} and doesn't emphasize partitions of equal 
sizes as opposed to for example spectral clustering and 
k-means\cite{von2007}.
\caption{Mirror Match with Clustering Algorithm (\emph{MMC})}
\label{alg-mmc}
\Require $images$ : set of images, $t \in \mathbb{R}$
\State $M\gets \varnothing$
\State $F\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features
	\State $f_i\gets getFeatures(I_i)$
	\State $F\gets F \cup f_i$
\State $A\gets getAdjecencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $A_{norm}\gets 1 - normalize(A)$
\State $A_{pruned}\gets pruneEdges(A_{norm},\alpha)$
\State $P\gets cluster(A_{pruned})$
\ForAll{$p \in P$} \Comment p is a set of feature points
	\State $M\gets M \cup getMatches(p, t, F)$
\Return matches
While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping around 2.5\% 
of edges seems to work well. Figure \ref{fig:graph} shows the result of 
clustering the feature points displayed as a graph.
\begin{SCfigure*}
	\makebox[0.7\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\includegraphics[width=\textwidth]{images/MMC_graph_full}
			\caption{Full Graph}
			\label{fig:full_graph}
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.3\textwidth}
			\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
			\caption{Cropped Section of Graph}
			\label{fig:cropped_graph}
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{SCfigure*}
The partitions group feature points by similarity which means that 
repetitive structures such as buildings often feature in larger 
partitions as exemplified in figure \ref{fig:compare_mirror}. %
\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{0.24\textwidth}
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.24\textwidth}
	\\
	\makebox[0.5\textwidth][c]{
		%~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)

		\begin{subfigure}[t]{0.5\textwidth}
			\caption{A Partition Example}
	\caption{Example of a partition of keypoints in an image pair. The 
		partition shown contains window corners from a building in the 
		left image and similar corners on the shirt in the right image.  
	\label{fig:compare_mirror}
\end{figure}
%The matching algorithm for \emph{MMC} (algorithm \ref{alg-getmatches}) 
%goes through the partitions one by one and selects matches in the 
%following way: If the partition has no edges going from a feature point 
%in one image to a feature point in another image, the partition is 
%discarded. If we have more than one edge going between images then 
%$getMatches$ will behave as \emph{MM} and make sure that a 
%correspondence is symmetric, that it has a ratio above the desired 
%threshold and that it isn't matching two feature points from the same 
%image. Since feature points are all guaranteed to be fairly similar 
%this means that we can set the ratio threshold much higher than in 
%\emph{MM} and include matches that would normally have been discarded 
%for not being sufficiently unique.
The matching for \emph{MMC} \emph{getMatches} finds matches within all 
partitions with more than two elements as done in \emph{MM}. However as 
can be seen in the example in figure \ref{fig:cropped_graph} many of the 
partitions contain only two feature points from different images linked 
by one edge. In this case we compare the ratio of this match to the next 
best match across all partitions.
%\begin{algorithm}
%\caption{Impl. of getMatches (\emph{from MMC algorithm})}
%\label{alg-getmatches}
%\begin{algorithmic}
%\Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
%all features
%\State $matches \gets \varnothing$
%\State $edges \gets \left\{getSimilarity(f_i, f_j) \mid getImg(f_i) \neq 
%getImg(f_j) \wedge f_i, f_j \in p \right\}$
%\If{$\left\vert edges \right\vert = 1$} \Comment If $\exists$ one edge 
%between images
%	\State $f_i,f_j \gets getInterImageFeatures(edges_{inter}, p)$
%	\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, features ~ 
%\backslash ~ \left\{f_i\right\})$
%	\State $f_s,f_t \gets getTwoNearestNeighbors(f_j, features ~ 
%\backslash ~ \left\{f_j\right\})$
%	\State $ratio_i \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_j)$
%	\State $ratio_j \gets getSimilarity(f_j, f_t) / getSimilarity(f_j, 
%f_i)$
%	\If{$ratio_i < t\wedge ratio_j < t$}
%		\State $matches \gets matches \cup (f_i, f_j)$
%	\EndIf
%\ElsIf{$\left\vert edges \right\vert > 1$} \Comment If $\exists$ more 
%than one edge between images
%	\ForAll{$f_i \in p$}
%		\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, p ~ \backslash 
%~ \left\{f_i\right\})$
%		\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_m)$
%		\If{$getImg(f_i) \neq getImg(f_m) \wedge ratio < t
%\wedge (f_m, f_i) \not\in matches$}
%			\State $matches \gets matches \cup (f_i, f_m)$
%		\EndIf
%	\EndFor
%\EndIf
%\Return matches
%\end{algorithmic}
%\end{algorithm}
\section{Experiments}
\label{experiment}
To reliably measure the accuracy of a matching method on real images we 
either need a set of image pairs tied by a homography or manually count 
the number of inliers. The latter becomes unfeasible the moment we 
attempt to test a non-trivial set of images. In 
\cite{mikolajczyk2005performance} Mikolajczyk and Schmid introduces a 
set of images to compare the performance of feature detectors. The test 
set covers different types of image variance such as different lighting, 
blur, rotation as well as viewpoint change.However in practice the view 
point change has gone on to be used the most for later tests given that 
most of the other test pairs have since become easy to 
match\footnote{see \cite{wu2011robust} or \cite{delponte2006svd} for 
examples}.
%Inspired by the graffiti image we have collected a set of image pairs 
%featuring murals of a few different artists\footnote{Including Banksy, 
%Blu and Shepard Fairey among others}. Each pair consist of images of 
%the same motive taken from different angles, often including repetitive 
%patterns and texture. Based on this image set the two algorithms are 
%tested on a series of image patches from each pair. To verify that the 
%algorithms proposed are reliable we need to test that we get good 
%matches when images overlap and that we get no matches when they don't.  
%This means that it isn't enough just testing against pairs of images 
%that we know match.  We also need to test against pairs that might look 
%like they could match but in fact don't.
To increase the amount of images that feature view port changes we have 
collected a set of $8$ image pairs featuring murals from different 
angles.  From these pairs we can create test suites by cropping out 
$250$ pixel by $250$ pixel patches of the original pair with a random 
vertical and horizontal offset. Given a source image pair we can obtain 
$n$ pairs of patches where the two patches in any pair might or might 
not overlap.  This ensures that the patch pairs whith no overlap still 
retain a basic similarity to each other, while the patches that do 
overlap often only overlap on a small part of their surface. This 
dataset will from here on be referred to as the \emph{Murals} dataset.  
Figure \ref{fig:fairey} shows an example of possible pairs of test 
images produced from two source images to the left.  Producing $n$ such 
pairs allows us to test not just how well the matching algorithm 
performs on corresponding images, but also how many false positives we 
get on similar images that don't overlap.\footnote{The set of source 
	images and homographies as well as the script to generate the 
cropped test sets based on them have been made available online together 
with the exact test sets used in this paper on 
\href{https://github.com/arnfred/Murals}{www.github.com/arnfred/Murals}}.
%\begin{figure*}
%	\centering
%	\includegraphics[width=\textwidth]{images/murals}
%	\caption{Images in the mural test set}
%	\label{fig:mural}
%\end{figure*}
\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{0.044\textwidth}
			\includegraphics[width=\textwidth]{images/pair_example1}
			\label{fig:fairey1}
		~ %add desired spacing between images, e. g. ~, \quad, (or a 
		%blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.044\textwidth}
			\includegraphics[width=\textwidth]{images/pair_example2}
			\label{fig:fairey2}
		~ %
		\begin{subfigure}[t]{0.37\textwidth}
			\includegraphics[width=\textwidth]{images/crop_examples}
	\caption{Example of test set produced by a source pair of images}
	\label{fig:fairey}
\end{figure}
%\subsection{Experimental setup}
Using the image test sets described above, the \emph{MM} and \emph{MMC} 
algorithms have been tested against \emph{SIFT}, the algorithm Lowe 
proposed in \cite{lowe2004sift} as well as the \emph{Isodata} matching 
algorithm proposed in \cite{das2008event} which uses geometric 
information to improve the matching. For each test set of cropped image 
pairs the four algorithms\footnote{\emph{MM}, \emph{MMC}, \emph{SIFT} 
and \emph{Isodata}} are applied to each pair of images ($100$ per test 
set), and the threshold for each algorithm is varied over 50 evenly 
spaced values. For each match returned by any of the algorithms, the 
match is registered as an inlier if given $H$ relating $I_1$ with $I_2$ 
and a match ($p_1, p_2$) the match is an inlier if $\left\vert Hp_1 - 
p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 \right\vert < 5$. The 
experiments are run on the $8$ image pairs in the test set, as well as 
the graf image from \cite{mikolajczyk2005performance}.
\section{Results}
\label{results}
%\begin{figure}
%	\makebox[\textwidth][c]{%
%		\begin{subfigure}[t]{0.35\textwidth}
%			\centering
%			\includegraphics[width=\textwidth]{images/fairey_burma_1}
%			\caption{Burma by Fairey (1)}
%			\label{fig:burma1}
%		\end{subfigure}%
%		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
%		%(or a blank line to force the subfigure onto a new line)
%		\begin{subfigure}[t]{0.35\textwidth}
%			\centering
%			\includegraphics[width=\textwidth]{images/fairey_burma_2}
%			\caption{Burma by Fairey (2)}
%			\label{fig:burma2}
%		\end{subfigure}%
%		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
%		%(or a blank line to force the subfigure onto a new line)
%		\begin{subfigure}[t]{0.35\textwidth}
%			\centering
%			\includegraphics[width=\textwidth]{images/result_fairey_burma}
%			\caption{X: Nb of correct matches. Y: accuracy}
%			\label{fig:result_burma}
%		\end{subfigure}%
%	}%
%	\label{fig:burma}
%	\caption{results on Burma by Fairey}
%\end{figure}
%\begin{figure}
%	\makebox[\textwidth][c]{%
%		\begin{subfigure}[t]{0.25\textwidth}
%			\centering
%			\includegraphics[width=\textwidth]{images/fairey_lady_1}
%			\caption{Lady by Fairey (1)}
%			\label{fig:lady1}
%		\end{subfigure}%

>>>>

\documentclass[12pt,journal]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
\usepackage[font={small}]{caption}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\title{Reliable feature point matching without the use of geometric 
constraints}
%\author{Jonas Arnfred and what to put here exactly?}

@@

containing viewport changes and a simple method to generate a large 

@@

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing every 
feature point of one image with all feature points of the other and 
finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. Ranking the 
scores by their uniqueness and picking the $n$ best produces a set of 
correspondences that are distinctly matched across the two images. This 
method will be referred to as \emph{SIFT} in this paper.

In turn the two methods proposed in this paper are inspired by a simple 
but novel idea. If we have two images and a given feature point in the 
first image is better matched with other feature points from the 
\emph{same} image than points in the other image, then any matches of 
this feature point to points in the other image is deemed unreliable and 
discarded.  This approach carries no implicit assumptions about the 
geometric consistency of matches and can as such easily be extended with 
other geometric solutions when appropriate.

@@

\section{Matching Methods}

@@

{\fontsize{10}{10}\selectfont

@@

		\begin{subfigure}[t]{0.32\textwidth}
		\enspace %
		\begin{subfigure}[t]{0.32\textwidth}
		\enspace %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.32\textwidth}

@@

			\caption{\emph{SIFT}}

@@

		\begin{subfigure}[t]{0.32\textwidth}
		\enspace%
		\begin{subfigure}[t]{0.32\textwidth}

@@

		\enspace %add desired spacing between images, e. g. ~, \quad, 
		%\qquad (or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.32\textwidth}

@@

			\caption{\emph{MMC} Partition Example}

@@

matches that are found within the same image. }%
	\label{fig:comparemirror}%
\end{figure*}%

In figure \ref{fig:comparemirror} an example of the result of \emph{MM} 
is illustrated. Notice how the by the bottom of the image admits 
correspondences that can conveniently be removed as done in sub figure 
\ref{fig:without} where only the actual correspondences returned by the 
algorithm are shown.

\subsection{Mirror Match with Clustering (\emph{MMC})}
%
As opposed to \emph{MM}, \emph{MMC} diverges from traditional 
non-geometric feature matching by clustering the feature points by 
similarity. This process yields partitions of fairly similar feature 
points that we can match using the same approach as \emph{MM}. 
%
%Before we introduce the implementation details we will go over the 
%problem of graph clustering and how it relates to feature matching.
%
%\subsubsection{Graph Clustering}
%
%Given a set of feature points from two images, $im_1$ and $im_2$: $F_k = 
%\{k_i, k_j$ for $k_i \in im_1, k_j \in im_2\}$ as well as a matching 
%function $M(k_i, k_j) \rightarrow \mathbb{R}$ that takes two feature 
%points in an image and returns their matching score, we can define a 
%matrix $A$ where each element $A_ij = M(k_i, k_j)$. A can be interpreted 
%as the \emph{adjecency matrix} of the fully connected graph where each 
%vertex corresponds to a keypoint and the edge between two vertices has a 
%weight equal to the distance between the two corresponding keypoints.
%
%This representation reduces the problem of partitioning the keypoints 
%into groups to that of graph clustering or community structure depending 
%on the context. In the literature there are various ways of clustering a 
%graph according to different measures of what constitutes an optimal 
%partitioning. Traditionally the most used clustering algorithms have 
%been K-means and spectral clustering, but in recent years a host of new 
%algorithms have been proposed based on both Newman's concept of graph 
%modularity\footnote{Introduced in \cite{girvan2002}, discussed in 
%\cite{brandes2007} and used in \cite{blondel2008} as well as others} as 
%well as information theoretical measures\footnote{See for example 
%\cite{rosvall2008}} and the Potts spin model from physics\footnote{Used 
%in \cite{ronhovde2009}} just to mention a few approaches. Many of the 
%new algorithms differ from K-means clustering and Spectral clustering in 
%that they don't require the number of expected to clusters to be 
%specified beforehand\footnote{Among the aforementioned methods, this is 
%true for \cite{blondel2008} and \cite{rosvall2008}}.  Furthermore, on 
%tests done using randomly generated graphs with a known partitioning 
%\cite{blondel2008}, \cite{rosvall2008} and \cite{ronhovde2009} perform 
%markedly better than spectral clustering and 
%K-means\cite{lancichinetti2009}.
%
%The performance of clustering algorithms is a complicated issue, since 
%an optimal clustering given the same graph can vary depending on the 
%application. Spectral clustering for example will usually return a 
%partitioning where each partition is roughly equal in size\footnote{As 
%mentioned in \cite{von2007}} while the Louvain 
%clustering\cite{blondel2008} might return partitions of very uneven 
%size, even if the modularity measurement has been shown to penalize very 
%small clusters\cite{brandes2007}. Both behaviours can be beneficial 
%depending on the application, but when clustering feature points, 
%maintaining clusters of an even size usually means that some clusters 
%will be '\emph{catch-all}' clusters where the feature points that don't 
%fit anywhere else are grouped together. The necessity of specifying the 
%amount of partitions in for example Spectral clustering or Pott's model 
%clustering further exacerbates the issue since smaller partitions are 
%then combined into one to achieve the right amount of partitions.

We use the Louvain Method\cite{blondel2008} for clustering feature 
points since it's relatively fast and performs 
well\cite{lancichinetti2009}, doesn't require 
parameters\cite{blondel2008} and doesn't emphasize partitions of equal 
sizes as opposed to for example spectral clustering and 
k-means\cite{von2007}.
%
\begin{algorithm}
\caption{Mirror Match with Clustering Algorithm (\emph{MMC})}
\label{alg-mmc}
{\fontsize{10}{10}\selectfont
\begin{algorithmic}
\Require $images$ : set of images, $t \in \mathbb{R}$
\State $M\gets \varnothing$
\State $F\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features
	\State $f_i\gets getFeatures(I_i)$
	\State $F\gets F \cup f_i$
\EndFor
\State $A\gets getAdjecencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $A_{norm}\gets 1 - normalize(A)$
\State $A_{pruned}\gets pruneEdges(A_{norm},\alpha)$
\State $P\gets cluster(A_{pruned})$
\ForAll{$p \in P$} \Comment p is a set of feature points
	\State $M\gets M \cup getMatches(p, t, F)$
\EndFor \\
\Return matches
\end{algorithmic}
}
\end{algorithm}

While the louvain clustering algorithm doesn't require any parameters in 
itself, it tends towards clustering all feature points together in the 
same partition if the graph is very connected.  To ensure that the graph 
is well clustered, the adjacency matrix is pruned so only edges above a 
certain threshold is kept. From empirical analysis keeping around 2.5\% 
of edges seems to work well. Figure \ref{fig:graph} shows the result of 
clustering the feature points displayed as a graph.
%
\begin{SCfigure*}
	\makebox[0.7\textwidth][c]{%
		\begin{subfigure}[t]{0.35\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_full}
			\caption{Full Graph}
			\label{fig:full_graph}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.3\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/MMC_graph_cropped}
			\caption{Cropped Section of Graph}
			\label{fig:cropped_graph}
		\end{subfigure}%
	}%
	\caption{The partitioned feature graph. Each vertex represents a 
		feature point while an edge represent a high similarity between 
		points. A partition is a connected group marked by a color.  Due 
		to a limited amount of colors some disconnected partitions have 
		been assigned the same color. The edge color of each node 
		signifies which image it belongs to. In the detailed crop to the 
		right the various cluster sizes can be seen, ranging from 
	hundreds of feature points to only two or three.}
	\label{fig:graph}
\end{SCfigure*}

The partitions group feature points by similarity which means that 
repetitive structures such as buildings often feature in larger 
partitions as exemplified in figure \ref{fig:pitts_partition}.
%
%\begin{figure}
%	\makebox[0.5\textwidth][c]{%
%		\begin{subfigure}[t]{0.24\textwidth}
%			\centering
%			\includegraphics[width=\textwidth]{images/MMC_pitts_source}
%			\caption{Source image}
%			\label{fig:pitts_source}
%		\end{subfigure}%
%		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
%		%(or a blank line to force the subfigure onto a new line)
%	}%
%	\\
%	\makebox[0.5\textwidth][c]{
%		%~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
%		%(or a blank line to force the subfigure onto a new line)
%		
%		\begin{subfigure}[t]{0.5\textwidth}
%			\centering
%			\includegraphics[width=\textwidth]{images/MMC_partition}
%			\caption{A Partition Example}
%			\label{fig:pitts_partition}
%		\end{subfigure}%
%	}%
%	\caption{Example of a partition of keypoints in an image pair. The 
%		partition shown contains window corners from a building in the 
%		left image and similar corners on the shirt in the right image.  
%	}
%	\label{fig:compare_mirror}
%\end{figure}
%
%The matching algorithm for \emph{MMC} (algorithm \ref{alg-getmatches}) 
%goes through the partitions one by one and selects matches in the 
%following way: If the partition has no edges going from a feature point 
%in one image to a feature point in another image, the partition is 
%discarded. If we have more than one edge going between images then 
%$getMatches$ will behave as \emph{MM} and make sure that a 
%correspondence is symmetric, that it has a ratio above the desired 
%threshold and that it isn't matching two feature points from the same 
%image. Since feature points are all guaranteed to be fairly similar 
%this means that we can set the ratio threshold much higher than in 
%\emph{MM} and include matches that would normally have been discarded 
%for not being sufficiently unique.

The matching for \emph{MMC} \emph{getMatches} finds matches within all 
partitions with more than two elements as done in \emph{MM}. However as 
can be seen in the example in figure \ref{fig:cropped_graph} many of the 
partitions contain only two feature points from different images linked 
by one edge. In this case we compare the ratio of this match to the next 
best match across all partitions.
%\begin{algorithm}
%\caption{Impl. of getMatches (\emph{from MMC algorithm})}
%\label{alg-getmatches}
%\begin{algorithmic}
%\Require $p$ : set of features, $t\in \mathbb{R}$, $features$ : Set of 
%all features
%\State $matches \gets \varnothing$
%\State $edges \gets \left\{getSimilarity(f_i, f_j) \mid getImg(f_i) \neq 
%getImg(f_j) \wedge f_i, f_j \in p \right\}$
%\If{$\left\vert edges \right\vert = 1$} \Comment If $\exists$ one edge 
%between images
%	\State $f_i,f_j \gets getInterImageFeatures(edges_{inter}, p)$
%	\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, features ~ 
%\backslash ~ \left\{f_i\right\})$
%	\State $f_s,f_t \gets getTwoNearestNeighbors(f_j, features ~ 
%\backslash ~ \left\{f_j\right\})$
%	\State $ratio_i \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_j)$
%	\State $ratio_j \gets getSimilarity(f_j, f_t) / getSimilarity(f_j, 
%f_i)$
%	\If{$ratio_i < t\wedge ratio_j < t$}
%		\State $matches \gets matches \cup (f_i, f_j)$
%	\EndIf
%\ElsIf{$\left\vert edges \right\vert > 1$} \Comment If $\exists$ more 
%than one edge between images
%	\ForAll{$f_i \in p$}
%		\State $f_m,f_n \gets getTwoNearestNeighbors(f_i, p ~ \backslash 
%~ \left\{f_i\right\})$
%		\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
%f_m)$
%		\If{$getImg(f_i) \neq getImg(f_m) \wedge ratio < t
%\wedge (f_m, f_i) \not\in matches$}
%			\State $matches \gets matches \cup (f_i, f_m)$
%		\EndIf
%	\EndFor
%\EndIf
%
%\Return matches
%\end{algorithmic}
%\end{algorithm}
%
\section{Experiments}
\label{experiment}
%
To reliably measure the accuracy of a matching method on real images we 
either need a set of image pairs tied by a homography or manually count 
the number of inliers. The latter becomes unfeasible the moment we 
attempt to test a non-trivial set of images.

In \cite{mikolajczyk2005performance} Mikolajczyk and Schmid introduces a 
set of images to compare the performance of feature detectors. The test 
set covers different types of image variance such as different lighting, 
blur, rotation as well as viewpoint change. Inspired by this 
dataset\footnote{In particular by the 'Graf' image set} and motivated by 
the need for more image pairs featuring viewport changes, we have 
compiled a set of $8$ image pairs consisting of the same subject taken 
from two different angles. The images are collected from Flickr's 
database of images published under a creative commons 
licence\footnote{\href{http://creativecommons.org/}{www.creativecommons.org}, 
a copyleft license permitting redistribution} and feature murals which 
makes it possible to relate points in the image pairs with a homography.  
The images have been cropped to feature the same motive and resized to 
$900$px by $600$px. This dataset will be referred to as the 
\emph{Murals} dataset in this paper. The images can be seen in figure 
\ref{fig:murals}.

\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{images/murals}
	\caption{Images in the mural test set}
	\label{fig:murals}
\end{figure*}

%However in practice the view point change has gone on to be used the 
%most for later tests given that most of the other test pairs have since 
%become easy to match\footnote{see \cite{wu2011robust} or 
%\cite{delponte2006svd} for examples}.

%Inspired by the graffiti image we have collected a set of image pairs 
%featuring murals of a few different artists\footnote{Including Banksy, 
%Blu and Shepard Fairey among others}. Each pair consist of images of 
%the same motive taken from different angles, often including repetitive 
%patterns and texture. Based on this image set the two algorithms are 
%tested on a series of image patches from each pair. To verify that the 
%algorithms proposed are reliable we need to test that we get good 
%matches when images overlap and that we get no matches when they don't.  
%This means that it isn't enough just testing against pairs of images 
%that we know match.  We also need to test against pairs that might look 
%like they could match but in fact don't.

%
\begin{figure}
	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example1}
			\label{fig:fairey1}
		\end{subfigure}%
		\enspace %add desired spacing between images, e. g. ~, \quad, 
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.048\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/pair_example2}
			\label{fig:fairey2}
		\end{subfigure}%
		\enspace %
		\begin{subfigure}[t]{0.36\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/crop_examples}
		\end{subfigure}%
	}%
	\caption{Example of test set produced by a source pair of images}
	\label{fig:fairey}
\end{figure}
%
%\subsection{Experimental setup}
The \emph{MM} and \emph{MMC} algorithms have been tested against 
\emph{SIFT}, the algorithm Lowe proposed in \cite{lowe2004sift} as well 
as the \emph{Isodata} matching algorithm proposed in \cite{das2008event} 
which uses geometric information to improve the matching. The comparison 
has been done using the \emph{Murals} dataset, the \emph{Graf} set from 
\cite{mikolajczyk2005performance} and two images\footnote{100\_1942.jpg 
and 100\_1941.jpg} from \cite{gallagher2008} illustrated in figure 
\ref{fig:pitts_partition}.

Test sets have been generated from the image pairs by cropping out 
$250$px by $250$px patches with a random vertical and horizontal offset.  
Given a source image pair we produce $100$ pairs of patches where the 
two patches in any pair might or might not overlap.  This ensures that 
the patch pairs with no overlap still retain a basic similarity to each 
other, while the patches that do overlap often only overlap on a small 
part of their surface. Figure \ref{fig:fairey} shows an example of 
possible pairs of test images produced from a source image pair.  
Producing $n$ such pairs allows us to test not just how well the 
matching algorithm performs on a variety of overlap but also how many 
false positives we get on similar images that don't overlap at 
all\footnote{The set of source images and homographies as well as the 
	script to generate the cropped test sets based on them have been 
made available online together with the exact test sets used in this 
paper on 
\href{https://github.com/arnfred/Murals}{www.github.com/arnfred/Murals}}.

To judge whether a possible match ($p_1, p_2$) should be counted as an 
inlier given a homography $H$ relating two images $I_1$ and $I_2$ we 
test if $p_1$ and $p_2$ satisfies the following criterium:

$$\left\vert Hp_1 - p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 
\right\vert < 5$$
%
\section{Results}
\label{results}
